
## Assignment 1

## 0 Preparation step
# Create a folder called assignment 1:
```{r}
if(!file.exists("./assignment1")){dir.create("./assignment1")}
```

# Download and unzip file:
```{r}
fileurl <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip"
download.file(fileurl, destfile = "./assignment1/assignment.zip")
unzip(zipfile = "./assignment1/assignment.zip", exdir = "./assignment1")
```

# Read the file into R and convert Date and Time variable into Date and Time format, the remaining variable to numeric format
```{r}
library(readr)
setwd("D:/Statistics/R/R data/Exploratory_data_assignment1/assignment1")
assignment <- read.csv("household_power_consumption.txt", sep = ";", stringsAsFactors = FALSE) 
```

# Converting formate of Date and create new variable call DateTime in Assignment dataframe
```{r}
assignment$Date <- as.Date(assignment$Date, format = "%d/%m/%Y")
assignment$DateTime <- strptime(paste(assignment$Date,assignment$Time),format = "%Y-%m-%d %H:%M:%S")
```

# Choose the data from the dates 2007-02-01 and 2007-02-02
```{r}
assignment <- assignment[assignment$Date >= "2007-2-1" & assignment$Date <= "2007-2-2",]
```

# Convert the other variable into numeric class
```{r}
assignment[,3] <- as.numeric(assignment[,3])
assignment[,4] <- as.numeric(assignment[,4])
assignment[,5] <- as.numeric(assignment[,5])
assignment[,6] <- as.numeric(assignment[,6])
assignment[,7] <- as.numeric(assignment[,7])
assignment[,8] <- as.numeric(assignment[,8])
lapply(assignment,class)
```

# 1. Constructing plot 1 and save
```{r}
with(assignment, hist(assignment$Global_active_power, xlab = "Global Active Power(kilowatts)", col = "red", main = "Global Active Power"))

dev.copy(device = png, width = 480, height = 480, file = "Plot1.png")
```

# 2. Constructing plot 2 and save
```{r}
with(assignment, plot(DateTime, Global_active_power, type = "l", xlab = "", ylab = "Global Active Power (kilowatts)", xaxt = "n"))

axis(side = 1, at = c(as.numeric(assignment$DateTime[1]), as.numeric(assignment$DateTime[1441]), as.numeric(assignment$DateTime[2880])), labels = c("Thu","Fri","Sat"))

dev.copy(device = png, width = 480, height = 480, file = "Plot2.png")
```

# 3. Constructing plot 3 and save

```{r}
with(assignment, plot(DateTime, Sub_metering_1, type = "n", xaxt = "n", xlab = "", ylab = "Energy sub metering"))

axis(side = 1, at = c(as.numeric(assignment$DateTime[1]), as.numeric(assignment$DateTime[1441]), as.numeric(assignment$DateTime[2880])), labels = c("Thu","Fri","Sat"))

with(assignment, lines(DateTime, Sub_metering_1))

with(assignment, lines(DateTime, Sub_metering_2, col = "red"))

with(assignment, lines(DateTime, Sub_metering_3, col = "blue"))

legend("topright", lty = 1, col=c("black", "red", "blue"), legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))

dev.copy(device = png, width = 480, height = 480, file = "Plot3.png")      
```


# 4. Constructing plot 4 and save
# Creating cluster 2x2
```{r}
      par(mfrow = c(2,2))
```

# Creating top-left plot 
```{r}
      with(assignment, plot(DateTime, Global_active_power, type = "l", xlab = "", ylab = "Global Active Power", xaxt = "n"))
      axis(side = 1, at = c(as.numeric(assignment$DateTime[1]),
                            as.numeric(assignment$DateTime[1441]),
                            as.numeric(assignment$DateTime[2880])),labels = c("Thu","Fri","Sat"))
```

# Creating top-right plot
```{r}
 with(assignment, plot(DateTime, Voltage, type = "l", xlab = "datetime", ylab = "Voltage", xaxt = "n"))
      axis(side = 1, at = c(as.numeric(assignment$DateTime[1]),
                            as.numeric(assignment$DateTime[1441]),
                            as.numeric(assignment$DateTime[2880])),labels = c("Thu","Fri","Sat"))
```

# Creating bottom-left plot
```{r}
 with(assignment, plot(DateTime, Sub_metering_1, type = "n", xaxt = "n",xlab = "", ylab = "Energy sub metering"))

axis(side = 1, at = c(as.numeric(assignment$DateTime[1]),as.numeric(assignment$DateTime[1441]),as.numeric(assignment$DateTime[2880])),labels = c("Thu","Fri","Sat"))

with(assignment, lines(DateTime, Sub_metering_1))

with(assignment, lines(DateTime, Sub_metering_2, col = "red"))

with(assignment, lines(DateTime, Sub_metering_3, col = "blue"))

legend("topright",lty = 1, col = c("black","red","blue"),legend = c("Sub_metering_1","Sub_metering_2","Sub_metering_3"))
```

# Creating bottom right plot
```{r}
with(assignment, plot(DateTime, Global_reactive_power, type = "l", xlab = "datetime",ylab = "Global_reactive_power", xaxt = "n"))

axis(side = 1, at = c(as.numeric(assignment$DateTime[1]),as.numeric(assignment$DateTime[1441]),as.numeric(assignment$DateTime[2880])),labels = c("Thu","Fri","Sat"))
```

# Saving plot
```{r}
dev.copy(device = png, width = 480, height = 480, file = "Plot4.png")      

```

# End
```{r}
dev.off()
```
